Mission-time LTL (MLTL) Formula Validation Via Regular Expressions\thanks{Work supported in part by [omitted for blind review] {\bf Note that all references to appendices will be replaced with URLs in the final paper; they are appendices only for blind review.}}

### 2022-10-14 paper/Paper.tex
many format edits,turned correctness proof and time complexity into theorems, more edits to correctness proof, etc - Chiara, Jeremy, Laura, Zili

We structure the paper around our contributions as follows. In section \ref{prereq}, we build on the semantics of MLTL to define a computation, and its bit string representation. % of a computation that is used throughout the paper and the WEST program. Finally, we describe negation normal form (NNF), a standardized way to express MLTL formulas. By writing the MLTL formulas in NNF, the WEST program is able to run much more efficiently.
%
In section \ref{regex}, we recursively define regular expressions encapsulating the satisfying computations of MLTL formulas and prove that our new translation algorithm is both sound and complete. In addition, we provide a calculation for the minimum computation length required to describe all the satisfying computations of a MLTL formula that slightly improves upon existing calculations in the literature. Finally, we show an application of the regular expressions of the satisfying computations by using them to prove an MLTL theorem. In particular, we show that $\phi\mathcal{U}_{[a,b]}(\phi\mathcal{U}_{[0,c]}\psi) = \phi\mathcal{U}_{[a,b+c]}\psi$ and $\phi\mathcal{R}_{[a,b]}(\phi\mathcal{R}_{[0,c]}\psi) = \phi\mathcal{R}_{[a,b+c]}\psi$.
%
We introduce a new tool, called WEST (acronym redacted for blind review), that implements automated validation in section \ref{specs} %, we give the WEST code specifications. In particular, we describe the grammar that the program follows and give an overview of the important functions in the program.
%
and calculate its space and time complexity, both theoretically and experimentally, in section \ref{complex}. We show that the theoretical worst-case complexity for both space and time is doubly exponential, but that the average case fares far better.
%
Section \ref{test} proves correctness of the algorithms and provides a test suite to show correctness of implementation with high confidence. Intelligent fuzzing techniques contribute to test suite construction from a state diagram representing the control flow of the algorithm. We also verify the correctness of outputs of the WEST program against a na\"ive brute force implementation.
%
Section \ref{simpsection}, provides a combinatorial theorem for simplifying certain outputs of the WEST program. Formulas of the form $\phi \equiv \psi$ to check for equivalence between two expressions can lead to outputs that are non-trivial to simplify. Section \ref{Conclusion} discusses impacts and future work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.1in}
\section{Preliminaries: Mission-time LTL and Bit String Computations} \label{prereq}
\vspace{-0.1in}
%\subsection{Mission-time Linear Temporal Logic}
{\bf Mission-time Linear Temporal Logic (MLTL)} \cite{LVR22} is a finite variation of LTL over bounded, closed, discrete intervals of the form $[a,b]$ where $a,b \in \mathbb{N}$ and $0 \leq a \leq b$. The syntax of an MLTL formula $\phi$ over a set of atomic propositions $\mathcal{AP}$, where $p \in \mathcal{AP}$ is a propositional variable, is given by the following BNF grammar:%
\vspace{-0.1in}
$$\phi := \top \ | \ \bot \ | \ p \ | \ \neg \psi \ | \ \psi_1 \land \psi_2 \ | \ \psi_1 \lor \psi_2 \ | \ \mathcal{F}_{[a,b]} \psi \ | \ \mathcal{G}_{[a,b]} \psi \ | \ \psi_1 \mathcal{U}_{[a,b]} \psi_2 \ | \ \psi_1 \mathcal{R}_{[a,b]} \psi_2 \footnote{For simplicity, we do not include parentheses in the grammar, but the WEST program requires parentheses (see Section \ref{specs}). We encode Release directly rather than as the dual of Until \cite{bmc,LVR22}.}$$
\vspace{-0.1in}
%
\begin{definition}[\emph{Computation}]
A computation $\pi$ of length $m$ is a sequence $\{\pi[i]\}_{i = 0}^{m-1}$ of sets of propositional variables, $\pi[i] \subseteq \mathcal{AP}$, where the $i^{th}$ set holds the propositional variables that are satisfied at the $i^{th}$ time step. That is, a propositional variable $p$ is true at time step $i$ if and only if $p \in \pi[i]$. We denote the suffix of $\pi$ starting at $i$ (including $i$) by $\pi_i$. Note that $\pi_0 = \pi$.
\end{definition}

@@

To describe the bit string representations of the satisfying computations for a given MLTL formula, we use the alphabet $\Sigma =$ \{`$0$' , `$1$', `$,$'\} and define $S$ as an abbreviation for $(0 \ | \ 1)$.
Let $p_0, p_2, ..., p_{n-1}$ be propositional variables, $\phi, \psi$ well-formed MLTL formulas in negation normal form. We define the regular expression of satisfying computations for an MLTL formula as follows:\\
%
%
% ZILI: Failed attempt to get formulas side by side
% \begin{minipage}{0.45 \textwidth}
%     \begin{align}
%     &\text{reg}(\top) = S^n\\
%     &\text{reg}(\bot) = \emptyset \\
%     &\text{reg}(p_k) = S^{k}1S^{n-k-1} \\
%     &\text{reg}(\neg p_k) = S^{k}0S^{n-k-1} \\
%     &\text{reg}(\phi \lor \psi) =  \text{reg}(\phi) \lor \text{reg}(\psi)\\
%     &\text{reg}(\phi \land \psi) = \text{reg}(\phi) \land \text{reg}(\psi)
%     \end{align}
% \end{minipage}
% \begin{minipage}{0.55 \textwidth}
%     \begin{align}
%     &\text{reg}(\mathcal{F}_{[a,b]} \phi) = \bigvee_{i=a}^{b} (S^n,)^i \text{reg}(\phi) \\
%     &\text{reg}(\mathcal{G}_{[a,b]} \phi) = \bigwedge_{i=a}^{b} (S^n,)^i \text{reg}(\phi)\\
%     &\text{reg}(\phi \ \mathcal{U}_{[a,b]} \psi) =  \bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i-1]}\phi \land \mathcal{G}_{[i, i]} \psi\right) \\
%     &\text{reg}(\phi \mathcal{R}_{[a,b]} \psi) =  \text{reg}\left(\mathcal{G}_{[a,b]}\psi\right) \lor \bigvee_{i=a}^{b-1} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right) \label{release}
%     \end{align}
% \end{minipage}
%
%ORIGINAL:
%% \begin{align}
%%      &\text{reg}(\top) = S^n\\
%%      &\text{reg}(\bot) = \emptyset \\
%%      &\text{reg}(p_k) = S^{k}1S^{n-k-1} \\
%%      &\text{reg}(\neg p_k) = S^{k}0S^{n-k-1} \\
%%      &\text{reg}(\phi \lor \psi) =  \text{reg}(\phi) \lor \text{reg}(\psi)\\
%%      &\text{reg}(\phi \land \psi) = \text{reg}(\phi) \land \text{reg}(\psi) \\
%%      &\text{reg}(\mathcal{F}_{[a,b]} \phi) = \bigvee_{i=a}^{b} (S^n,)^i \text{reg}(\phi) \\
%%      &\text{reg}(\mathcal{G}_{[a,b]} \phi) = \bigwedge_{i=a}^{b} (S^n,)^i \text{reg}(\phi)\\
%%      &\text{reg}(\phi \ \mathcal{U}_{[a,b]} \psi) =  \bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i-1]}\phi \land \mathcal{G}_{[i, i]} \psi\right) \\
%%      &\text{reg}(\phi \mathcal{R}_{[a,b]} \psi) =  \text{reg}\left(\mathcal{G}_{[a,b]}\psi\right) \lor \bigvee_{i=a}^{b-1} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right) \label{release}
%% \end{align}
%%
     $\text{reg}(\top) = S^n$ \hfill 
     $\text{reg}(\bot) = \emptyset$ \\
     $\text{reg}(p_k) = S^{k}1S^{n-k-1}$ \hfill
     $\text{reg}(\neg p_k) = S^{k}0S^{n-k-1}$ \\
     $\text{reg}(\phi \lor \psi) =  \text{reg}(\phi) \lor \text{reg}(\psi)$\hfill
     $\text{reg}(\phi \land \psi) = \text{reg}(\phi) \land \text{reg}(\psi)$ \\
     $\text{reg}(\mathcal{F}_{[a,b]} \phi) = \bigvee_{i=a}^{b} (S^n,)^i \text{reg}(\phi)$ \hfill
     $\text{reg}(\mathcal{G}_{[a,b]} \phi) = \bigwedge_{i=a}^{b} (S^n,)^i \text{reg}(\phi)$\\
     $\text{reg}(\phi \ \mathcal{U}_{[a,b]} \psi) =  \bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i-1]}\phi \land \mathcal{G}_{[i, i]} \psi\right)$ \\
     $\text{reg}(\phi \mathcal{R}_{[a,b]} \psi) =  \text{reg}\left(\mathcal{G}_{[a,b]}\psi\right) \lor \bigvee_{i=a}^{b-1} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right)$ %\label{release} %KYR: I notice this label was not used...

%     \vspace{-0.1in}
%     \subsection{Minimum Computation Length}
%          \vspace{-0.1in}
 %These regular expressions describe all the satisfying computations of an MLTL formula of computation length $\text{cplen}(\phi)$, which is defined recursively below:

@@

We define the computation length $\text{cplen}(\phi)$ of an MLTL formula $\phi$ recursively:\\

@@

\indent \qquad   $\text{cplen}(p_k) = \text{cplen}(\neg p_k) = 1$\\
\indent \qquad   $\text{cplen}(\phi \land \psi) = \text{cplen}(\phi \lor \psi) = \max(\text{cplen}(\phi), \text{cplen}(\psi))$\\
\indent \qquad   $\text{cplen}(\mathcal{G}_{[a,b]} \phi) = \text{cplen}(\mathcal{F}_{[a,b]} \phi) = b + \text{cplen}(\phi)$\\
\indent \qquad   $\text{cplen}(\phi \mathcal{U}_{[a,b]} \psi) = \text{cplen}(\phi \mathcal{R}_{[a,b]} \psi) = b + \max(\text{cplen}(\phi)-1, \text{cplen}(\psi))$

@@

    Let $0 \leq a\leq b \in \mathbb{N}$ and let $\phi, \psi$ be well-formed MLTL formulas in NNF. The minimum computation length of Until and Release is given by $\text{cplen}(\phi \mathcal{U}_{[a,b]} \psi) = \text{cplen}(\phi \mathcal{R}_{[a,b]} \psi) = b + \max(\text{cplen}(\phi)-1, \text{cplen}(\psi))$.

@@

 We write $\mathscr{L}(\text{reg}(\phi))$ to denote the language of reg($\phi$). That is, the set of computations represented by the regular expression $\text{reg}(\phi)$. % The proof of Theorem \ref{SoundComplete} appears in Appendix \ref{Appendix-SoundComplete}.

 \begin{theorem}[\emph{Soundness and Completeness}] \label{SoundComplete}
 For any well formed MLTL formula $\phi$ in negation normal form, a computation $\pi$ with $|\pi| = \text{cplen}(\phi)$ satisfies $\phi$ and  if and only if $\pi \in \mathscr{L}(\emph{reg}(\phi))$.
 \end{theorem}

%KYR: this totally goes here in the journal version but in an appendix/on a website for the conference version
%%  \begin{proof}
%%  We proceed by induction on the length of the formula.\\

%%  \noindent \textbf{Base Cases.}\\
%%  \noindent\textbf{reg($\top$)}\\
%%  Any computation $\pi$ satisfies $\top$, and $\mathscr{L}(\text{reg}(\top))$ is the set of all computations of one time step. Padding conventions extends this to the set of all computations of any positive number of time steps, so we are done.\\

%%  \noindent\textbf{reg($\bot$)}\\
%%  There are no computations that satisfy $\bot$ and $\mathscr{L}(\text{reg}(\bot))$ is the empty string, so we are done.\\

%%  \noindent\textbf{reg($p_k$)}\\
%%  For $0\leq k \leq n-1$, consider the propositional variable $p_k$. If $\pi$ satisfies $p_k$, then $p_k$ evaluates to true at $\pi[0]$. This is equivalent to writing that $\pi[0]$ is of the form 
%%  $$S^{k} 1 S^{n-k-1}$$
%%  which is precisely reg($p_k$).\\

%%  \noindent\textbf{reg($\neg p_k$)}\\
%%  If $\pi$ satisfies $\neg p_k$, then $\neg p_k$ evaluates to true at $\pi[0]$. This is equivalent to writing that $\pi[0]$ is of the form 
%%  $$S^{k} 0 S^{n-k-1}$$
%%  which is precisely reg($\neg p_k$).\\

%%  \noindent\textbf{Inductive Step.}\\
%%  Suppose the theorem holds for MLTL formulas $\varphi$ and $\psi$. We now show that it holds for $\varphi \land \psi$, $\varphi \lor \psi$, $\mathcal{F}_{[a,b]} \varphi$, $\mathcal{G}_{[a,b]} \varphi$, $\varphi \ \mathcal{U}_{[a,b]} \psi$, and $\varphi \mathcal{R}_{[a,b]} \psi$.\\

%%  \noindent \textbf{reg($\varphi \land \psi$)}\\
%%  We know that $\pi \vDash \varphi \land \psi \text{ iff } \pi \vDash \varphi \text{ and } \pi \vDash \psi$.\\
%%  By the inductive hypothesis, $\pi \vDash \varphi$ iff $\pi \in \mathscr{L}(\text{reg}(\varphi))$ and 
%%  \\$\pi \vDash \psi$ iff $\pi \in \mathscr{L}(\text{reg}(\psi))$, so
%%  $$\pi \vDash \varphi \land \psi \text{ iff } \pi \in (\mathscr{L}(\text{reg}(\varphi)) \land \mathscr{L}(\text{reg}(\psi))) = \mathscr{L}(\text{reg}(\varphi \land \psi)).$$

%%  \noindent \textbf{reg($\varphi \lor \psi$)}\\
%%  We know that $\pi \vDash \varphi \lor \psi \text{ iff } \pi \vDash \varphi \text{ or } \pi \vDash \psi$.\\
%%  By the inductive hypothesis, $\pi \vDash \varphi$ iff $\pi \in \mathscr{L}(\text{reg}(\varphi))$ and 
%%  \\$\pi \vDash \psi$ iff $\pi \in \mathscr{L}(\text{reg}(\psi))$, so
%%  $$\pi \vDash \varphi \lor \psi \text{ iff } \pi \in (\mathscr{L}(\text{reg}(\varphi)) \lor \mathscr{L}(\text{reg}(\psi))) = \mathscr{L}(\text{reg}(\varphi \lor \psi)).$$

%%  \noindent \textbf{reg($\mathcal{F}_{[a,b]} \varphi$)}\\
%%  We know that $\pi \vDash \mathcal{F}_{[a,b]} \varphi \text{ iff } |\pi| > a \text{ and } \exists i \in [a,b] \text{ such that } \pi_i \vDash \varphi$.\\
%%  If $|\pi| = \text{cplen}(\mathcal{F}_{[a,b]} \varphi)$, $|\pi| > a$. Likewise, $\pi \in \mathscr{L}\left(\text{reg}\left(\mathcal{F}_{[a,b]} \varphi \right)\right)$ implies $|\pi| = \text{cplen}(\mathcal{F}_{[a,b]} \varphi)$, so the length condition is satisfied.\\
%%  By the inductive hypothesis, $\pi_i \vDash \varphi$ iff $\pi_i \in \mathscr{L}(\text{reg}(\varphi))$, so $\pi \in \mathscr{L}((S^n,)^i \text{reg}(\varphi))$ for some $i \in [a,b]$. Equivalently, 
%%  $$\pi \in \mathscr{L}\left(\bigvee_{i=a}^{b} (S^n,)^i \text{reg}(\phi)(,S^n)^{b-i}\right) = \mathscr{L}(\text{reg}(\mathcal{F}_{[a,b]}\varphi)).$$

%%  \noindent \textbf{reg($\mathcal{G}_{[a,b]} \varphi$)}\\
%%  We know that $\pi \vDash \mathcal{G}_{[a,b]}\varphi \text{ iff } |\pi| \leq a \text{ or } \forall i \in [a,b] \ \pi_i \vDash \varphi$.\\
%%  Since $|\pi| = \text{cplen}(\mathcal{G}_{[a,b]} \varphi)$ and $\text{cplen}(\mathcal{G}_{[a,b]} \varphi) > a$, the first option for satisfying the formula never occurs.\\
%%  By the inductive hypothesis, $\pi_i \vDash \varphi$ iff $\pi_i \in \mathscr{L}(\text{reg}(\varphi))$, so $\pi \in \mathscr{L}((S^n,)^i \text{reg}(\varphi))$ for all $i \in [a,b]$. Equivalently, 
%%  $$\pi \in \mathscr{L}\left(\bigwedge_{i=a}^{b} (S^n,)^i \text{reg}(\phi)(,S^n)^{b-i}\right) = \mathscr{L}(\text{reg}(\mathcal{F}_{[a,b]}\varphi)).$$

%%  \noindent \textbf{reg($\varphi \ \mathcal{U}_{[a,b]} \psi$)}\\
%%   We have that 
%%   $$\pi \vDash \varphi \ \mathcal{U}_{[a,b]} \psi \text{ iff } |\pi| > a \text{ and } \exists i \in [a,b] \text{ such that } (\pi_i \vDash \psi \text{ and } \forall a \leq j < i, \pi_j \vDash \varphi).$$
%%   As argued in the Finally case, the length condition is satisfied.\\
%%   By the inductive hypothesis, $\pi_i \vDash \psi$ iff $\pi_i \in \mathscr{L}(\text{reg}(\psi))$, or equivalently, \\
%%   $\pi \in \mathscr{L}(\text{reg}(\mathcal{G}_{[i,i]} \psi))$. Also, $\pi_j \vDash \varphi$ if and only if $\pi_j \in \mathscr{L}(\text{reg}(\varphi))$.\\
%%   By the argument used in the Global case, we see that $\forall a \leq j < i, \ \pi_j \vDash \varphi$ is equivalent to $\pi \in \mathscr{L}(\text{reg}(\mathcal{G}_{[a, i-1]} \varphi))$.\\
%%   Thus $\pi \vDash \varphi \ \mathcal{U}_{[a,b]} \psi$ iff $\pi \in \mathscr{L}((\text{reg}(\mathcal{G}_{[i,i]} \psi) \land \text{reg}( \mathcal{G}_{[a, i-1]} \varphi)))$ for some $i \in [a,b]$, or equivalently, 
%%   $$\pi \in \mathscr{L}\left(\bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i-1]}\phi \land \mathcal{G}_{[i, i]} \psi\right)\right) = \mathscr{L}\left(\text{reg}(\varphi \ \mathcal{U}_{[a,b]} \psi)\right).$$

%%   \noindent \textbf{reg($\varphi \mathcal{R}_{[a,b]} \psi$)}\\
%%   We have that
%%   \begin{align*}
%%     &\pi \vDash \varphi \mathcal{R}_{[a,b]} \psi \text{ iff } |\pi| \leq a \text{ or } \forall i \in [a,b] \ \pi_i \vDash \psi \text{ or } \\ 
%%     &\exists j\in[a,b] \text{ such that }
%%     (\pi_j \vDash \varphi \text{ and } \forall a\leq k \leq j, \pi_k \vDash \psi).
%%   \end{align*}
%%   As argued in the Global case, the first option for satisfying the formula never occurs.\\
%%   By the Global case, the statement $\forall i\in[a,b] \ \pi_i \vDash \psi$ is equivalent to $\pi \in \mathscr{L}(\text{reg}(\mathcal{G}_{[a,b]} \psi))$ and, by the Finally case. the statement $\exists j\in[a,b]$ such that ($\pi_j \vDash \varphi$ and $\forall a\leq k \leq j$, $\pi_k \vDash \psi$) is equivalent to $\pi \in \mathscr{L}\left(\bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right)\right)$. Hence

%%   \begin{align*}
%%     \pi \vDash \varphi \mathcal{R}_{[a,b]} \psi \text{ iff } \pi &\in \mathscr{L}\left(\text{reg}(\mathcal{G}_{[a,b]} \psi) \lor \bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right)\right) \\
%%     &= \mathscr{L}\left(\text{reg}(\mathcal{G}_{[a,b]} \psi) \lor \bigvee_{i=a}^{b-1} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right)\right)\\
%%     &= \mathscr{L}\left(\text{reg}(\varphi \mathcal{R}_{[a,b]} \psi)\right).
%%   \end{align*}

%%  \noindent This completes the inductive step, and thus the proof. Since this proof addresses all possible MLTL formulas in negation normal form, it shows completeness along with soundness.
%%  \end{proof}
% \newpage
%\vspace{-0,1in}
%\subsection{Nested Until and Release Rewriting Theorem}
%\vspace{-0,1in}
 We prove the correctness of our formula rewriting algorithm using our regular expressions from Definition \ref{TRE}. As detailed in Section \ref{complex}, nested binary temporal operators create the worst-case time and space complexity for the WEST program. Although the WEST program does not implement this theorem by default, we would expect selected applications of rewriting to improve runtime. We use this theorem to simplify MLTL formulas, making them easier to interpret.
% Theorem \ref{NestedUR}'s proof appears in Appendix \ref{Appendix-NestedUR}.

@@

\section{WEST Algorithm and Analysis} \label{WESTalg}
We now present the  full WEST algorithm, which recursively computes all satisfying temporal regular expressions to a given MLTL formula (in NNF). 
For convenience, we use sets to represent alternation of regular expressions; for $n$ regular expressions $t_0, ..., t_{n-1}$, we'll write $\{t_1, t_2, ..., t_n\} = \bigcup_{i=1}^n t_i$. 
Additionally, we provide details for performing set intersection of temporal regular expressions, and the algorithm for the temporal operator $\mathcal{F}$. 
Note how reg\_U parallels the regular expression defined for $\mathcal{F}$, and the algorithms for the other three temporal operators follow an identical structure.

@@

% ZILI: Presenting algorithm vertically instead
\hspace{-5mm}
\begin{minipage}{0.52\textwidth} 
\begin{algorithm}[H] 
\caption{WEST Algorithm \\
Inputs: $\phi$ - MLTL formula in NNF \\
$\phi_1$ and $\phi_2$ below are subformulas of $\phi$\\
$n$ - number of propositional variables \\
Output: set of REGEX satisfying $\phi$}
\begin{algorithmic}[1]
    \Procedure{reg}{string $\phi$, int $n$}
        \If{$\phi$ is $\top$ or $\bot$}
            \State return reg\_prop\_const($\phi$, n)
        \EndIf

@@

For time complexity, through a similar analysis, we've found that the time-complexity of all of the above functions is the same as their space complexity. In particular, the time complexities of $\phi = p_k, \scriptstyle{\sim}\textstyle p_k, \top, \bot$, as well as that of set intersection, \textit{join}, and the operator $\mathcal{U}$ match their respective space complexities. Thus the worst case time complexity is $O(\mathcal{C}_U^{\delta ^ \ell} \cdot (\ell + 1)^{\delta^{\ell + 1}})$ as well. 
If in the worst case no simplification occurs in any call of \textit{set\_intersect}, space complexity remains unchanged, but simplifying a set of regular expressions is cubic in input size. In practice however, both time and space complexities are much more optimistic than worst-case estimates.

@@

For the first simulation, we consider 2 iterations, 5 propositional variables, 10 for DELTA, and 10 for INTERVAL\_MAX. We obtain the following plots:\\

@@

    \label{InputLengthVsOutputLength}

@@

    \label{InputLengthVsRuntime}

@@

    \label{fig:InVsOutLen}

@@

    \label{fig:InLenVsRunTime}

@@

\Procedure{reg\_F}{set reg\_phi, int $a$, int $b$, int $n$, bool simp} 

>>>>

We structure the paper around our contributions as follows. In section \ref{prereq}, we build on the semantics of MLTL to define a computation, and its bit string representation. % of a computation that is used throughout the paper and the WEST program. Finally, we describe negation normal form (NNF), a standardized way to express MLTL formulas. By writing the MLTL formulas in NNF, the WEST program is able to run much more efficiently.
%
In section \ref{regex}, we recursively define regular expressions encapsulating the satisfying computations of MLTL formulas and prove that our new translation algorithm is both sound and complete. In addition, we provide a calculation for the minimum computation length required to describe all the satisfying computations of a MLTL formula that slightly improves upon existing calculations in the literature. Finally, we show an application of the regular expressions of the satisfying computations by using them to prove an MLTL theorem. In particular, we show that $\phi\mathcal{U}_{[a,b]}(\phi\mathcal{U}_{[0,c]}\psi) = \phi\mathcal{U}_{[a,b+c]}\psi$ and $\phi\mathcal{R}_{[a,b]}(\phi\mathcal{R}_{[0,c]}\psi) = \phi\mathcal{R}_{[a,b+c]}\psi$.
%
We introduce a new tool, called WEST (acronym redacted for blind review), that implements automated validation in section \ref{WESTalg} %, we give the WEST code specifications. In particular, we describe the grammar that the program follows and give an overview of the important functions in the program.
%
and calculate its space and time complexity, both theoretically and experimentally. We show that the theoretical worst-case complexity for both space and time is doubly exponential, but that the average case fares far better.
%
Section \ref{test} proves correctness of the algorithms and provides a test suite to show correctness of implementation with high confidence. Intelligent fuzzing techniques contribute to test suite construction from a state diagram representing the control flow of the algorithm. We also verify the correctness of outputs of the WEST program against a na\"ive brute force implementation.
%
Section \ref{simpsection}, provides a combinatorial theorem for simplifying certain outputs of the WEST program. Formulas of the form $\phi \equiv \psi$ to check for equivalence between two expressions can lead to outputs that are non-trivial to simplify. Section \ref{Conclusion} discusses impacts and future work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.1in}
\section{Preliminaries: Mission-time LTL and Bit String Computations} \label{prereq}
\vspace{-0.1in}
%\subsection{Mission-time Linear Temporal Logic}
{\bf Mission-time Linear Temporal Logic (MLTL)} \cite{LVR22} is a finite variation of LTL over bounded, closed, discrete intervals of the form $[a,b]$ where $a,b \in \mathbb{N}$ and $0 \leq a \leq b$. The syntax of an MLTL formula $\phi$ over a set of atomic propositions $\mathcal{AP}$, where $p \in \mathcal{AP}$ is a propositional variable, is given by the following BNF grammar:%
\vspace{-0.1in}
$$\phi := \top \ | \ \bot \ | \ p \ | \ \neg \psi \ | \ \psi_1 \land \psi_2 \ | \ \psi_1 \lor \psi_2 \ | \ \mathcal{F}_{[a,b]} \psi \ | \ \mathcal{G}_{[a,b]} \psi \ | \ \psi_1 \mathcal{U}_{[a,b]} \psi_2 \ | \ \psi_1 \mathcal{R}_{[a,b]} \psi_2 \footnote{For simplicity, we do not include parentheses in the grammar, but the WEST program requires parentheses (see Section \ref{WESTalg}). We encode Release directly rather than as the dual of Until \cite{bmc,LVR22}.}$$
\vspace{-0.1in}
%
\begin{definition}[\emph{Computation}]
A computation $\pi$ of length $m$ is a sequence $\{\pi[i]\}_{i = 0}^{m-1}$ of sets of propositional variables, $\pi[i] \subseteq \mathcal{AP}$, where the $i^{th}$ set holds the propositional variables that are satisfied at the $i^{th}$ time step. That is, a propositional variable $p$ is true at time step $i$ if and only if $p \in \pi[i]$. We denote the suffix of $\pi$ starting at $i$ (including $i$) by $\pi_i$. Note that $\pi_0 = \pi$.
\end{definition}

@@

To describe the bit string representations of the satisfying computations for a given MLTL formula, we use the alphabet $\Sigma =$ \{``$0$'' , ``$1$'', ``$,$''\} and define $S$ as an abbreviation for $(0 \ | \ 1)$.
Let $p_0, p_1, ..., p_{n-1}$ be propositional variables, $\phi, \psi$ well-formed MLTL formulas in negation normal form. We define the regular expression of satisfying computations for an MLTL formula as follows:\\

% ZILI: Left Aligned Version
\vspace{-7mm}
\hspace{-8.5mm}
\begin{minipage}{0.45 \textwidth}
    \begin{align}
    &\text{reg}(\top) = S^n\\
    &\text{reg}(p_k) = S^{k}1S^{n-k-1} \\
    &\text{reg}(\phi \lor \psi) =  \text{reg}(\phi) \lor \text{reg}(\psi)\\
    &\text{reg}(\mathcal{G}_{[a,b]} \phi) = \bigwedge_{i=a}^{b} (S^n,)^i \text{reg}(\phi)
    \end{align}
\begin{minipage}{0.55\textwidth}
    \begin{align}
    &\text{reg}(\bot) = \emptyset \\
    &\text{reg}(\neg p_k) = S^{k}0S^{n-k-1} \\
    &\text{reg}(\phi \land \psi) = \text{reg}(\phi) \land \text{reg}(\psi) \\
    &\text{reg}(\mathcal{F}_{[a,b]} \phi) = \bigvee_{i=a}^{b} (S^n,)^i \text{reg}(\phi)
    \end{align}
\vspace{-5mm}\\
\begin{minipage}{0.99\textwidth}
\begin{align}
&\text{reg}(\phi \ \mathcal{U}_{[a,b]} \psi) =  \bigvee_{i=a}^{b} \text{reg}\left(\mathcal{G}_{[a,i-1]}\phi \land \mathcal{G}_{[i, i]} \psi\right)\\
\vspace*{-10mm}&\text{reg}(\phi \mathcal{R}_{[a,b]} \psi) =  \text{reg}\left(\mathcal{G}_{[a,b]}\psi\right) \lor \bigvee_{i=a}^{b-1} \text{reg}\left(\mathcal{G}_{[a,i]}\psi \land \mathcal{G}_{[i, i]} \phi\right) \label{release}
\end{align}

@@

We define the computation length $\emph{cplen}(\phi)$ of an MLTL formula $\phi$ recursively:\\

@@

\indent \qquad   $\emph{cplen}(p_k) = \emph{cplen}(\neg p_k) = 1$\\
\indent \qquad   $\emph{cplen}(\phi \land \psi) = \emph{cplen}(\phi \lor \psi) = \max(\emph{cplen}(\phi), \emph{cplen}(\psi))$\\
\indent \qquad   $\emph{cplen}(\mathcal{G}_{[a,b]} \phi) = \emph{cplen}(\mathcal{F}_{[a,b]} \phi) = b + \emph{cplen}(\phi)$\\
\indent \qquad   $\emph{cplen}(\phi \mathcal{U}_{[a,b]} \psi) = \emph{cplen}(\phi \mathcal{R}_{[a,b]} \psi) = b + \max(\emph{cplen}(\phi)-1, \emph{cplen}(\psi))$

@@

    Let $0 \leq a\leq b \in \mathbb{N}$ and let $\phi, \psi$ be well-formed MLTL formulas in NNF. The minimum computation length of Until and Release is given by $\emph{cplen}(\phi \mathcal{U}_{[a,b]} \psi) = \emph{cplen}(\phi \mathcal{R}_{[a,b]} \psi) = b + \max(\emph{cplen}(\phi)-1, \emph{cplen}(\psi))$.

@@

 We write $\mathscr{L}(\text{reg}(\phi))$ to denote the language of reg($\phi$). That is, the set of computations represented by the regular expression $\text{reg}(\phi)$.
 \begin{theorem}[\emph{Soundness and Completeness}] \label{SoundComplete}
 For any well formed MLTL formula $\phi$ in negation normal form, a computation $\pi$ with $|\pi| = \emph{cplen}(\phi)$ satisfies $\phi$ and  if and only if $\pi \in \mathscr{L}(\emph{reg}(\phi))$.
 We prove the correctness of our formula rewriting algorithm using our regular expressions from Definition \ref{TRE}. As detailed in Section \ref{WESTalg}, nested binary temporal operators create the worst-case time and space complexity for the WEST program. Although the WEST program does not implement this theorem by default, we would expect selected applications of rewriting to improve runtime. We use this theorem to simplify MLTL formulas, making them easier to interpret.
% Theorem \ref{NestedUR}'s proof appears in Appendix \ref{Appendix-NestedUR}.

@@

\section{WEST Algorithm and Analysis} \label{WESTalg}
We now present the  full WEST algorithm, which recursively computes all satisfying temporal regular expressions to a given MLTL formula (in NNF). 
For convenience, we use sets to represent alternation of regular expressions; for $n$ regular expressions $t_0, ..., t_{n-1}$, we'll write $\{t_0, ..., t_{n-1}\} = \bigcup_{i=0}^{n-1} t_i$. 
Additionally, we provide details for performing set intersection of temporal regular expressions, and the algorithm for the temporal operator $\mathcal{F}$. 
Note how reg\_U parallels the regular expression defined for $\mathcal{F}$, and the algorithms for the other three temporal operators follow an identical structure.

@@

% ZILI: Presenting algorithm vertically instead
\hspace{-5mm}
\begin{minipage}{0.52\textwidth} 
\begin{algorithm}[H] 
\caption{WEST Algorithm}
Inputs: $\phi$ - MLTL formula in NNF \\
$\phi_1$ and $\phi_2$ below are subformulas of $\phi$\\
$n$ - number of propositional variables \\
Output: set of REGEX satisfying $\phi$
\begin{algorithmic}[1]
    \Procedure{reg}{string $\phi$, int $n$}
        \If{$\phi$ is $\top$ or $\bot$}
            \State return reg\_prop\_const($\phi$, n)
        \EndIf

@@

For time complexity, through a similar analysis, we've found that the time-complexity of all of the above functions is the same as their space complexity. In particular, the time complexities of $\phi = p_k, \scriptstyle{\sim}\textstyle p_k, \top, \bot$, as well as that of set intersection, \textit{join}, and the operator $\mathcal{U}$ match their respective space complexities.
\begin{theorem}[\emph{Time Complexity}]
Given a well formed MLTL formula $\phi$, $\text{\emph{reg}}(\phi)$ has worst-case time complexity that is $O(\mathcal{C}_U^{\delta ^ \ell} \cdot (\ell + 1)^{\delta^{\ell + 1}})$, where $\ell$ is the number of logical connectives ($\land, \lor, \mathcal{F}, \mathcal{G}, \mathcal{R}, \mathcal{U}$) in $\phi$. 
\end{theorem}
If in the worst case no simplification occurs in any call of \textit{set\_intersect}, space complexity remains unchanged, but simplifying a vector of regular expressions is cubic in input size. In practice however, both time and space complexities are much more optimistic than worst-case estimates.

@@

For the first simulation, we consider 2 iterations, 5 propositional variables, 10 for DELTA, and 10 for INTERVAL\_MAX. We obtain plots \ref{InVsOutputLen} and \ref{InLenVsRunTime}:\\

@@

    \label{InVsOutputLen}

@@

    \label{InLenVsRunTime}

@@

    \label{fig:InVsOutLen2}

@@

    \label{fig:InLenVsRunTime2}

@@

\Procedure{reg\_F}{set $r_\phi$, int $a$, int $b$, int $n$, bool simp} 
